### 2022-10-03 manuscript/manuscript.tex
more intro notes

- Disentangling evolutionary processes from the patterns of genetic variation
across the genome. The structure of what can be under selection shapes these
patterns when selection acts; when drift acts, these patterns should not exist.

@@

@@

% could be in intro too:

As genomes are transmitted through the generations, lifetime events selectively
alter what is passed to the next generation. While selection truly acts on the
genetic variants that directly cause heritable fitness differences among
individuals, neighboring regions experience an indirect contraction in genetic
variability. This byproduct is known as \emph{linked selection}, and it occurs
when selection changes a site's frequency on timescale faster than
recombination can decouple that site from its genetic background. Because
linked selection modifies the genetic variability in a regionally-correlated
way, it creates patterns in genomic data that are used to understand the types
of selection processes acting on genomes.

Much work has been done on developing mathematical theory for two primary
selection processes: hitchhiking and background selection. In turn, statistical
methods have use this theory for a variety of wide-ranging empirical problems.
Early on, researchers found correlations between recombination rates and
genetic diversity in small genomes with limited recombination like
\emph{Drosophila} (XXX), consistent with linked selection processes acting on
genome-wide scales. With population genomic data, small-scale (XXX kb) dips in
genetic diversity was evidence of rare but strong selective sweeps in humans
(XXX) and other species (XXX).

models of background selection can
closely predict levels of human-chimp divergence at the megabase scale (XXX).

humans (XXX), though these offered only a partial picture of the selection
processes acting (XXX).

Linked
selection studies have found additional strong selective sweeps in taxa across
the kingdom of life (XXX), and attention turned towards estimating the relative
contributions of the hitchhiking and background processes (Eyalshiv).
processes (Eyalshiv).

From an evolutionary perspective, models of linked selection

>>>>

\section*{Introduction}

As genomes are transmitted through the generations, lifetime events selectively
alter what is passed to the next generation. While selection directly acts on
the genetic variants that cause heritable fitness differences among
individuals, neighboring regions experience an indirect contraction in genetic
variability. This byproduct is known as \emph{linked selection}, and it occurs
when selection changes a site's frequency on timescale faster than
recombination can decouple that site from its genetic background. Because
linked selection modifies the genetic variability in a regionally-correlated
way, it creates observable patterns in genomic data that are used to understand
the type of selection processes acting on genomes and estimate key evolutionary
quantities.

Much work has been done on developing mathematical theory for two linked
selection processes: hitchhiking and background selection. What distinguishes
these are the evolutionary processes that generate fitness variation, and the
resulting selection dynamics. Classic hitchhiking models assume new beneficial
variants occur and sweep to fixation (XXX). Theory predicts that classic sweeps
affect diversity on small spatial scales, with characteristic length
$\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
whether hitchhiking alone could explain the observed genome-wide correlation
between recombination rate and genetic diversity in some species (XXX). Since
the sweeps leave relatively narrow footprint, they influence genome-wide
diversity only when beneficial mutations are recurring and recombination is
relatively limited, as in \emph{Drosophila}.

By contrast, background selection occurs because of the steady influx of new
deleterious mutations into conserved regions each generation. Since far more
new mutations are deleterious than beneficial in functional regions, it would
seem like background selection would have a dominant role in shaping genetic
diversity. However, deleterious mutations segregate on a timescale inversely
proportional to their selection cost (XXX). Strongly deleterious mutations
eliminate neighboring genetic variability on the few chromosomes they may
recombine onto before the eventual extinction, and consequently have very weak
effects on larger spatial scales. Moderately weakly selected mutations can
persist for much longer periods, allowing for more recombination before their
eventual loss and produce reductions in diversity on shorter spatial scales.
The reduction in diversity under background selection is determined
approximately by $\nicefrac{U}{R}$, where $U = \mu_D L$ is the number of
deleterious mutations (at rate $\mu_D$ per basepair, per generation) in a
conserved region of $L$ basepairs per generation, and $R = r L$, which is the
total recombination per generation in Morgans. Since the deleterious mutation
rate is relatively constant across conserved regions, the reduction due to
background selection closely matches the composition of conserved regions and
recombination along the genome. Remarkably, simple mathematical models of
background selection closely predicts levels of human-chimp divergence at the
megabase scale (XXX) from genome-wide patterns of recombination and regions
conserved between species (McVicker).

This leads to some of the unidentifiability of hitchhiking versus background
selection.

and the reduction is determined
by

While the rate
new beneficial mutations enter the population is low (XXX),

The spatial scale of background selection is

kkkkkk

not all
adaptation occurs from such new beneficial mutations sweeping to fixation (XXX)

Other processes also act
on beneficial fitness variation to engender linked selection, such as polygenic
selection after an optimum shift \parencite{Buffalo2019-qs,Buffalo2020-my} or
soft sweeps from standing variation (XXX).

In turn, statistical methods have use this theory for a variety of wide-ranging
empirical problems.  Early on, researchers

While it is difficult to differentiate the selective processes
leading to linked selection, studies indicate background selection

hitchhiking events where a low-frequency

studies
in \emph{Drosophila} and humans indicate background selection has a large role
in shaping patterns of diversity, especially at large spatial scales (i.e.
megabases).

Another central empirical goal is to estimate the relative strengths of the
these two selection processes on levels of genetic diversity (Eyalshiv). Since
the effects of linked selection are determined in part by mutation rate and
distribution of fitness effects (DFE), these statistical methods also estimate
these important evolutionary parameters from genomic data. These estimates
should, in theory, align with others using different approaches, such as
trio-based estimates of the mutation rate and estimates of the DFE from
site-frequency spectra.

Understanding the extent to which linked selection influences other
evolutionary processes is a central question in evolutionary genetics. Two
central questions have emerged. First, while genetic drift is thought to be the
main source of random fluctuations in neutral allele frequencies, linked
selection can have a similar impact on neutral sites, increasing the variance
in allele frequency change as if the population size were effectively smaller.
Although linked selection causes significant reductions in genetic diversity in
some species, it is unlikely to explain the narrow range of observed diversity
across species \parencite{Buffalo2021-uf}. Second, by perturbing the
frequencies of neighboring alleles, selection can interfere with their
selection dynamics, known as Hill-Robertson interference (HRI, XXX). When
recombination is limited, simultaneous selection across the genome reduces the
odds beneficial mutations are fixed (HRI), effectively like selection was
occurring in an smaller population size. The Hill-Robertson effect is thought
to be weak in organisms in organisms with large recombination maps, perhaps
because of selection for increased recombination.

Our focus here is to begin to fix some deficiencies in the background selection
model. In particular,

% outline?
- Disentangling evolutionary processes from the patterns of genetic variation
across the genome. The structure of what can be under selection shapes these
patterns when selection acts; when drift acts, these patterns should not exist.

@@

\section*{Conclusion}

- care should be taken applying these methods to \emph{Drosophila}, since the
$\delta_{LD}$ term could be much larger.

@@

even when sites under selection are sparse.

From an evolutionary perspective, linked selection is important because by
modifying neighboring levels of genetic diversity, it is also directly altering
the selection dynamics at these neighboring sites.

Typically evolutionary genetic models consider
